\documentclass[12pt, a4paper]{report}
\usepackage[utf8]{inputenc}
\usepackage{graphicx}
\usepackage{hyperref}
\usepackage{listings}
\usepackage{color}
\usepackage{geometry}
\usepackage{float}
\usepackage{booktabs}
\usepackage{titlesec}
\usepackage{fancyhdr}

\geometry{margin=1in}

% Code highlighting settings
\definecolor{codegreen}{rgb}{0,0.6,0}
\definecolor{codegray}{rgb}{0.5,0.5,0.5}
\definecolor{codepurple}{rgb}{0.58,0,0.82}
\definecolor{backcolour}{rgb}{0.95,0.95,0.92}

\lstdefinestyle{mystyle}{
    backgroundcolor=\color{backcolour},   
    commentstyle=\color{codegreen},
    keywordstyle=\color{magenta},
    numberstyle=\tiny\color{codegray},
    stringstyle=\color{codepurple},
    basicstyle=\ttfamily\footnotesize,
    breakatwhitespace=false,         
    breaklines=true,                 
    captionpos=b,                    
    keepspaces=true,                 
    numbers=left,                    
    numbersep=5pt,                  
    showspaces=false,                
    showstringspaces=false,
    showtabs=false,                  
    tabsize=2
}
\lstset{style=mystyle}

\title{
    \vspace*{2cm}
    \textbf{\Huge FinanceIQ: Advanced Fintech Dashboard} \\
    \vspace{0.5cm}
    \Large Comprehensive Technical Documentation \\
    \vspace{2cm}
    \textbf{Project Report}
}
\author{Development Team}
\date{\today}

\begin{document}

\maketitle

\begin{abstract}
This document serves as a comprehensive technical report for \textbf{FinanceIQ}, an advanced personal finance management application. The system integrates a modern React-based frontend with a robust FastAPI backend to provide users with holistic financial insights. Key capabilities include real-time wealth management (stocks, gold, crypto), AI-powered financial advisory with LaTeX rendering, automated fraud detection using Isolation Forests, recurring payment tracking, and split-bill functionality. This report details the system architecture, database schema, implementation of key modules, and deployment instructions, intended for developers and stakeholders requiring a deep understanding of the project's inner workings.
\end{abstract}

\tableofcontents
\newpage

\chapter{Introduction}
\section{Project Overview}
FinanceIQ is designed to be a "Super App" for personal finance. It goes beyond simple expense tracking by incorporating intelligent features usually found in high-end wealth management platforms. It solves the problem of fragmented financial data by unifying expense tracking, investment portfolio management, and financial planning into a single, cohesive interface.

\section{Key Objectives}
\begin{itemize}
    \item \textbf{Unified Dashboard}: A "Bento Grid" style interface for at-a-glance metrics.
    \item \textbf{Smart Analytics}: ML-driven categorization and anomaly detection.
    \item \textbf{Wealth Tracking}: Real-time asset valuation via Yahoo Finance (yfinance).
    \item \textbf{AI Advisory}: Context-aware LLM integration for personalized advice.
    \item \textbf{Reporting}: Professional-grade PDF generation for audits and portfolios.
\end{itemize}

\chapter{System Architecture}
\section{Technology Stack}
The application follows a client-server architecture:

\subsection{Frontend}
\begin{itemize}
    \item \textbf{Framework}: React (Vite) for high-performance UI rendering.
    \item \textbf{Styling}: TailwindCSS for utility-first, responsive design.
    \item \textbf{State Management}: React Hooks (`useState`, `useEffect`).
    \item \textbf{Visualization}: Recharts for interactive charts (Area, Pie, Bar).
    \item \textbf{Data Fetching}: Axios for REST API communication.
    \item \textbf{Math Rendering}: `react-markdown` with `rehype-katex` for AI formulas.
    \item \textbf{Icons}: Lucide React for consistent visual language.
\end{itemize}

\subsection{Backend}
\begin{itemize}
    \item \textbf{Framework}: FastAPI (Python) for high-performance async endpoints.
    \item \textbf{Database}: SQLite (Development) / PostgreSQL (Production ready) with SQLAlchemy ORM.
    \item \textbf{Data Processing}: Pandas for analytics and CSV processing.
    \item \textbf{Machine Learning}: Scikit-learn (RandomForest, IsolationForest).
    \item \textbf{External APIs}: `yfinance` for market data, OpenRouter for LLM.
    \item \textbf{Reporting}: FPDF and ReportLab for PDF generation.
\end{itemize}

\chapter{Core Modules & Implementation}

\section{1. Authentication & Security}
The system uses JWT (JSON Web Tokens) for stateless authentication.
\begin{itemize}
    \item \textbf{Endpoints}: `/api/auth/register`, `/api/auth/token`.
    \item \textbf{Security}: Password hashing via `bcrypt` (Passlib).
    \item \textbf{Session}: Token storage in local storage / HTTP headers.
\end{itemize}

\section{2. The Unified Dashboard}
The dashboard serves as the command center.
\begin{itemize}
    \item \textbf{Layout}: Implements a responsive grid system.
    \item \textbf{Widgets}: Financial Health Score, Monthly Spend vs Budget, Asset Allocation, and Recent Transactions.
    \item \textbf{Interactivity}: Users can toggle time periods (Monthly, Yearly) and download reports directly from the dashboard.
\end{itemize}

\section{3. Wealth Management Engine}
This module tracks the user's Assets (Stocks, Gold, Crypto, Mutual Funds).
\subsection{Real-time Data Fetching}
Using the `yfinance` library, the backend fetches dynamic data:
\begin{lstlisting}[language=Python]
ticker = yf.Ticker(asset.symbol)
hist = ticker.history(period="1d")
current_price = float(hist['Close'].iloc[-1])
\end{lstlisting}
\subsection{Portfolio Reporting}
A sophisticated PDF generator (`reports.py`) creates Landscape-oriented portfolio reports.
\begin{itemize}
    \item \textbf{P\&L Calculation}: Real-time computation of Unrealized Gains/Losses.
    \item \textbf{Visuals}: Color-coded metrics (Green/Red arrows).
    \item \textbf{History}: A dedicated page tracking asset purchase history.
\end{itemize}

\section{4. AI Financial Advisor}
The "Brain" of the application.
\begin{itemize}
    \item \textbf{Context Injection}: The backend aggregates recent transactions, balance, and spending trends and feeds them into the system prompt.
    \item \textbf{LaTeX Support}: The LLM is instructed to output mathematical formulas in LaTeX format.
    \item \textbf{Rendering}: The frontend uses `ReactMarkdown` + `rehype-katex` to render complex equations (e.g., compound interest formulas).
\end{itemize}

\section{5. Fraud Detection System}
An integrated Machine Learning pipeline (`train_models.py`) runs in the background.
\begin{itemize}
    \item \textbf{Algorithm}: Isolation Forest (Unsupervised Learning).
    \item \textbf{Features}: Transaction Amount, Frequency, Weekend Spending, Category Z-Scores.
    \item \textbf{Response}: Anomalous transactions are flagged in high-priority red alerts on the "Fraud & Risk" page.
\end{itemize}

\section{6. Recurring Payments & Subscriptions}
Automatically detects and tracks repeating expenses.
\begin{itemize}
    \item \textbf{Logic}: Scans transaction history for similar amounts/descriptions on periodic intervals.
    \item \textbf{UI}: dedicated card view to cancel or monitor subscriptions.
\end{itemize}

\section{7. Splitwise Integration (Group Expenses)}
Allows users to manage shared expenses.
\begin{itemize}
    \item \textbf{Models}: `ExpenseGroup`, `GroupMember`, `SharedExpense`.
    \item \textbf{Logic}: Simplifies debts within a group context.
\end{itemize}

\chapter{Database Schema}
The application uses a Relational Database managed by SQLAlchemy.

\section{Key Tables}
\begin{itemize}
    \item \textbf{User}: `id`, `username`, `password_hash`.
    \item \textbf{Transaction}: `id`, `user_id`, `amount`, `category`, `date`, `description`.
    \item \textbf{Asset}: `id`, `user_id`, `symbol`, `quantity`, `purchase_price`, `purchase_date`, `type`.
    \item \textbf{Budget}: `id`, `user_id`, `category`, `limit`.
    \item \textbf{RecurringPayment}: `id`, `name`, `amount`, `frequency`.
    \item \textbf{ChatMessage}: `id`, `session_id`, `role`, `content` (Stores chat history).
\end{itemize}

\chapter{Installation & Deployment}
\section{Prerequisites}
\begin{itemize}
    \item Python 3.9+
    \item Node.js 18+
    \item SQLite (Pre-installed)
\end{itemize}

\section{Backend Setup}
\begin{lstlisting}[language=bash]
cd backend
python -m venv venv
source venv/bin/activate
pip install -r requirements.txt
python main.py
\end{lstlisting}

\section{Frontend Setup}
\begin{lstlisting}[language=bash]
cd frontend
npm install
npm run dev
\end{lstlisting}

\chapter{Conclusion}
FinanceIQ represents a significant leap in personal finance tooling. By combining traditional ledger capabilities with modern AI and real-time market data, it empowers users to take control of their financial destiny with precision and ease.

\end{document}
